---
title: "Convergence Clubs of wage in 34 Indonesian Provinces"
output: html_notebook
---

#Activate necessary libraries
```{r}
library(ConvergenceClubs)
library(tidyverse)
library(janitor)
library(tibble)
library(data.table)
library(ExPanDaR)
```


#Part I: Nominal Wage using definition "A" of worker
#Load the data
```{r}
library(readxl)
nominal.wage <- read_excel("WS Labour_new.xlsx", 
    sheet = "nominal_wage_A")
nominal.wage
```

#Select data only from 2008 onwards and remove "Region" column, and remove the observation of Indonesia
```{r}
nominal.wage <- nominal.wage %>% 
  select(-c(1,3:5)) %>% 
  subset(Province!="National")
```

#transform the wage series to log
```{r}
nominal.wage.log <- log(nominal.wage[,-1])
nominal.wage.log
```

#Remove short-run noise using HP filter
```{r}
f.nominal.wage.log <- apply(nominal.wage.log, 1, 
function(x){mFilter::hpfilter(x, freq=6.25, type="lambda")$trend} ) 
f.nominal.wage.log <- data.frame(Province = nominal.wage[,1], t(f.nominal.wage.log), stringsAsFactors=FALSE ) 
colnames(f.nominal.wage.log) <- colnames(nominal.wage) 
```

#Inspect the filtered data
```{r}
head(f.nominal.wage.log)
```

#Run the log-t-test
```{r}
H.nominal <- computeH(f.nominal.wage.log[,-1], quantity = "H")
round(estimateMod(H.nominal, time_trim=0.333, HACmethod = "FQSB"), 3)
```

#find convergence clubs
```{r}
clubs.nominal <- findClubs(f.nominal.wage.log, dataCols=2:14, unit_names = 1, refCol=14,
time_trim=0.333, cstar=0, HACmethod = 'FQSB')
```

```{r}
summary(clubs.nominal)
```

```{r}
print(clubs.nominal)
```

```{r}
plot(clubs.nominal)
```

```{r}
plot(clubs.nominal, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

#Merge clubs
```{r}
mclubs.nominal <- mergeClubs(clubs.nominal, mergeMethod='PS')
summary(mclubs.nominal)
```

```{r}
mclubs.nominal
```

```{r}
plot(mclubs.nominal)
```

```{r}
plot(mclubs.nominal, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

--------------
#Part II: Real Wage using definition "A" of worker
#Load the data
```{r}
library(readxl)
real.wage <- read_excel("WS Labour_new.xlsx", 
    sheet = "real_wage_A")
real.wage
```

#Select data only from 2008 onwards and remove "Region" column, and remove the observation of Indonesia
```{r}
real.wage <- real.wage %>% 
  select(-c(1,3:5)) %>% 
  subset(Province!="National")
```

#transform the wage series to log
```{r}
real.wage.log <- log(real.wage[,-1])
real.wage.log
```

#Remove short-run noise using HP filter
```{r}
f.real.wage.log <- apply(real.wage.log, 1, 
function(x){mFilter::hpfilter(x, freq=6.25, type="lambda")$trend} ) 
f.real.wage.log <- data.frame(Province = real.wage[,1], t(f.real.wage.log), stringsAsFactors=FALSE ) 
colnames(f.real.wage.log) <- colnames(real.wage) 
```

#Inspect the filtered data
```{r}
head(f.real.wage.log)
```

#Run the log-t-test
```{r}
H.real <- computeH(f.real.wage.log[,-1], quantity = "H")
round(estimateMod(H.real, time_trim=0.333, HACmethod = "FQSB"), 3)
```

#find convergence clubs
```{r}
clubs.real <- findClubs(f.real.wage.log, dataCols=2:14, unit_names = 1, refCol=14,
time_trim=0.333, cstar=0, HACmethod = 'FQSB')
```

```{r}
summary(clubs.real)
```

```{r}
print(clubs.real)
```

```{r}
plot(clubs.real)
```

```{r}
plot(clubs.real, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

#Merge clubs
```{r}
mclubs.real <- mergeClubs(clubs.real, mergeMethod='PS')
summary(mclubs.real)
```

```{r}
mclubs.real
```

```{r}
plot(mclubs.real)
```

```{r}
plot(mclubs.real, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

```{r}
class(mclubs.nominal)
```

```{r}
table.mclubs.nom <- map(mclubs.nominal, as.data.table)
df.mclubs.nom <- rbindlist(table.mclubs.nom, fill = T, idcol = T)
df.mclubs.nom
```
```{r}
colnames(df.mclubs.nom)[5] <- "Province"
df.mclubs.nom
```


```{r}
nominal.wage.club <- as.data.frame(
  inner_join(f.nominal.wage.log, df.mclubs.nom, by="Province") %>% 
  select(-contains(c("id","model")))
  ) 
nominal.wage.club
```

#Generate relative value for all clubs
```{r}
relative <- list()
for(a in 2:14) {
  relative[[a]] <- data.frame(nominal.wage.club$Province, nominal.wage.club[,a]/mean(nominal.wage.club[,a]))
colnames(relative[[a]])[1]<- "Province"
colnames(relative[[a]])[2]<- 
paste("rel",colnames(nominal.wage.club)[a],sep="_")
}

for (x in 2:13) {
  relative[[x+1]]<-left_join(relative[[x]],relative[[x+1]], by="Province")
}
```

```{r}
table.relative <- map(relative, as.data.table)
df.relative <- rbindlist(table.relative, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative <- join(df.relative,nominal.wage.club,by=c('Province')) %>% 
  select(Province,contains("rel"),clubs)
  
df.relative <- as.data.frame(df.relative)
```

```{r}
colnames(df.relative) <- gsub("rel_","",colnames(df.relative))
df.relative
```

```{r}
df.relative.long <- df.relative %>%  pivot_longer(-c(Province,clubs), names_to = "Time", values_to="Rel_Wage")
df.relative.long
```

```{r}
df.relative.path <- aggregate(Rel_Wage ~ clubs + Time, df.relative.long, mean) %>% 
  arrange(clubs)
```

```{r}
path_all <- df.relative.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=clubs, col=clubs)) + geom_line() +
  labs(title = "Transition path of all clubs") +
  theme_bw()
path_all
ggsave("path_all.png", width = 9, height = 5)
```

#Generate relative value for within club
#Create dataframe for each club
```{r}
club1 <- nominal.wage.club %>% 
  filter(clubs == "club1")
```

```{r}
relative.club1 <- list()
for(a in 2:14) {
  relative.club1[[a]] <- data.frame(club1$Province, club1[,a]/mean(club1[,a]))
colnames(relative.club1[[a]])[1]<- "Province"
colnames(relative.club1[[a]])[2]<- 
paste("rel",colnames(club1)[a],sep="_")
}

for (x in 2:13) {
  relative.club1[[x+1]]<-left_join(relative.club1[[x]],relative.club1[[x+1]], by="Province")
}
```


```{r}
table.relative.club1 <- map(relative.club1, as.data.table)
df.relative.club1 <- rbindlist(table.relative.club1, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.club1 <- join(df.relative.club1,club1,by=c('Province')) %>% 
  select(Province,contains("rel"),clubs)
  
df.relative.club1
```

```{r}
colnames(df.relative.club1) <- gsub("rel_","",colnames(df.relative.club1))
df.relative.club1
```

```{r}
df.relative.club1.long <- df.relative.club1 %>%  pivot_longer(-c(Province,clubs), names_to = "Time", values_to="Rel_Wage")
df.relative.club1.long
```

```{r}
df.relative.club1.path <- aggregate(Rel_Wage ~ Province + Time, df.relative.club1.long, mean) %>% 
  arrange(Province)
```

```{r}
path_club1 <- df.relative.club1.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Province, col=Province)) + geom_line() +
  labs(title = "Transition path of regions in Club 1") +
  theme_bw() +
   theme(legend.title = element_text(size = 8), 
               legend.text = element_text(size = 8))
path_club1
ggsave("path_club1.png", width = 9, height = 5)
```


```{r}
club2 <- nominal.wage.club %>% 
  filter(clubs == "club2")
```

```{r}
relative.club2 <- list()
for(a in 2:14) {
  relative.club2[[a]] <- data.frame(club2$Province, club2[,a]/mean(club2[,a]))
colnames(relative.club2[[a]])[1]<- "Province"
colnames(relative.club2[[a]])[2]<- 
paste("rel",colnames(club2)[a],sep="_")
}

for (x in 2:13) {
  relative.club2[[x+1]]<-left_join(relative.club2[[x]],relative.club2[[x+1]], by="Province")
}
```


```{r}
table.relative.club2 <- map(relative.club2, as.data.table)
df.relative.club2 <- rbindlist(table.relative.club2, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.club2 <- join(df.relative.club2,club2,by=c('Province')) %>% 
  select(Province,contains("rel"),clubs)
  
df.relative.club2
```

```{r}
colnames(df.relative.club2) <- gsub("rel_","",colnames(df.relative.club2))
df.relative.club2
```

```{r}
df.relative.club2.long <- df.relative.club2 %>%  pivot_longer(-c(Province,clubs), names_to = "Time", values_to="Rel_Wage")
df.relative.club2.long
```

```{r}
df.relative.club2.path <- aggregate(Rel_Wage ~ Province + Time, df.relative.club2.long, mean) %>% 
  arrange(Province)
```

```{r}
path_club2 <- df.relative.club2.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Province, col=Province)) + geom_line() +
  labs(title = "Transition path of regions in Club 2") +
  theme_bw() +
  theme(legend.title = element_text(size = 7), 
               legend.text = element_text(size = 7))
path_club2
ggsave("path_club2.png", width = 9, height = 5)
```

#Part II: Real Wage using definition "A" of worker
#Load the data
```{r}
library(readxl)
real.wage <- read_excel("WS Labour_new.xlsx", 
    sheet = "real_wage_A")
real.wage
```

#Select data only from 2008 onwards and remove "Region" column, and remove the observation of Indonesia
```{r}
real.wage <- real.wage %>% 
  select(-c(1,3:5)) %>% 
  subset(Province!="National")
```


#Transform to long format for ploting
```{r}
real.wage.long <- real.wage %>% 
  pivot_longer(!Province, names_to="Year", values_to="Wage")
```

```{r}
real.wage.long <- as.data.frame(real.wage.long)
real.wage.long$Year <- as.Date(as.factor(real.wage.long$Year), format = "%Y")
```

```{r}
real.wage.long$Year <- as.factor(as.numeric(real.wage.long$Year))
```


```{r}
prepare_trend_graph(real.wage.long, "Year", c("Wage"))$plot
```

```{r}
prepare_quantile_trend_graph(real.wage.long, "Year", c(0.05, 0.25, 0.5, 0.75, 0.95), "Wage")$plot +
  ggplot2::ylab("Wage (000 IDR)") + scale_x_continuous(breaks = 2008:2020) +
  theme_classic() + theme(axis.text.x=element_text(angle=90))
ggsave("wage_quantile.png", width = 9, height = 5)
```

```{r}
graph <- prepare_quantile_trend_graph(real.wage.long[c("Year", "Wage")], "Year", c(0.05, 0.25, 0.5, 0.75, 0.95))
graph$plot
```




#Plot wage of all provinces
```{r}
graph <- real.wage.long %>%
ggplot(aes(x = Year, y = Wage, group=Province)) +
geom_line(size=0.5, color = "dark gray") +
labs(x = "Time",
y = "Average real wage per month (thousand IDR)", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
graph
ggsave("wage_all.png", width = 9, height = 5)
```


```{r}
#install.packages("ggplotFL", repos="http://flr-project.org/R")
library(ggplotFL)
```


```{r}
G <- ggplot(real.wage.long, aes(x=Year, y=Wage)) +
  geom_flquantiles(probs=c(0.25, 0.50, 0.75), fill="red", alpha=0.25, show.legend=TRUE) + labs(x = "Time",
y = "Real wage", color="") + theme_classic()
G
ggsave("quantile.png")
```

```{r}
library(skimr)
```

```{r}
desc <- real.wage.long %>% 
  group_by(Province) %>% 
  skim() 
```

```{r}
desc <- desc %>% 
  select(!c(4:10))
```

```{r}
desc <- desc %>% 
 drop_na() %>% 
  write_excel_csv2("desc_stat.csv")
```



#Transform the wage series to log
```{r}
real.wage.log <- log(real.wage[,-1])
real.wage.log
```

#Remove short-run noise using HP filter
```{r}
f.real.wage.log <- apply(real.wage.log, 1, 
function(x){mFilter::hpfilter(x, freq=6.25, type="lambda")$trend} ) 
f.real.wage.log <- data.frame(Province = real.wage[,1], t(f.real.wage.log), stringsAsFactors=FALSE ) 
colnames(f.real.wage.log) <- colnames(real.wage) 
```

#Inspect the filtered data
```{r}
head(f.real.wage.log)
```

#Run the log-t-test
```{r}
H.real <- computeH(f.real.wage.log[,-1], quantity = "H")
round(estimateMod(H.real, time_trim=0.333, HACmethod = "FQSB"), 3)
```

#find convergence clubs
```{r}
clubs.real <- findClubs(f.real.wage.log, dataCols=2:14, unit_names = 1, refCol=14,
time_trim=0.333, cstar=0, HACmethod = 'FQSB')
```

```{r}
summary(clubs.real)
```

#We find three club convergence fo nominal wage


#Show the membership (provinces) of each club 

```{r}
print(clubs.real)
```

```{r}
plot(clubs.real)
```

```{r}
plot(clubs.real, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

#Merge clubs
```{r}
mclubs.real <- mergeClubs(clubs.real, mergeMethod='PS')
summary(mclubs.real)
```

```{r}
mclubs.real
```

```{r}
plot(mclubs.real)
```

```{r}
plot(mclubs.real, clubs=NULL, avgTP = TRUE, legend=TRUE)
```

#Convert clubs from list object to dataframe
```{r}
table.mclubs.real <- map(mclubs.real, as.data.table)
df.mclubs.real <- rbindlist(table.mclubs.real, fill = T, idcol = T)
df.mclubs.real
```

```{r}
colnames(df.mclubs.real)[c(1,5)] <- c("Club","Province")
df.mclubs.real
```

```{r}
real.wage.club <- as.data.frame(
  inner_join(f.real.wage.log, df.mclubs.real, by="Province") %>% 
  select(-contains(c("clubs","id","model")))
  ) 
real.wage.club
```

```{r}
real.wage.club$Club <- gsub("club", "Club ", real.wage.club$Club)
real.wage.club
```

#Transition path of all clubs
```{r}
relative.real <- list()
for(a in 2:14) {
  relative.real[[a]] <- data.frame(real.wage.club$Province, real.wage.club[,a]/mean(real.wage.club[,a]))
colnames(relative.real[[a]])[1]<- "Province"
colnames(relative.real[[a]])[2]<- 
paste("rel",colnames(real.wage.club)[a],sep="_")
}

for (x in 2:13) {
  relative.real[[x+1]]<-left_join(relative.real[[x]],relative.real[[x+1]], by="Province")
}
```

```{r}
table.relative.real <- map(relative.real, as.data.table)
df.relative.real <- rbindlist(table.relative.real, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.real <- join(df.relative.real,real.wage.club,by=c('Province')) %>% 
  select(Province,contains("rel"),Club)
  
df.relative.real <- as.data.frame(df.relative.real)
```

```{r}
colnames(df.relative.real) <- gsub("rel_","",colnames(df.relative.real))
df.relative.real
```

```{r}
df.relative.real.long <- df.relative.real %>%  pivot_longer(-c(Province,Club), names_to = "Time", values_to="Rel_Wage")
df.relative.real.long
```

```{r}
df.relative.real.path <- aggregate(Rel_Wage ~ Club + Time, df.relative.real.long, mean) %>% 
  arrange(Club)
```

```{r}
path_all.real <- df.relative.real.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Club, col=Club)) + geom_line() +
labs(x = "Time",
y = "Relative transition paths of all clubs", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
path_all.real
ggsave("path_all.real.png", width = 9, height = 5)
```

#Transition path of each club
#Club 1
```{r}
club1.real <- real.wage.club %>% 
  filter(Club == "Club 1")
```

```{r}
relative.club1.real <- list()
for(a in 2:14) {
  relative.club1.real[[a]] <- data.frame(club1.real$Province, club1.real[,a]/mean(club1.real[,a]))
colnames(relative.club1.real[[a]])[1]<- "Province"
colnames(relative.club1.real[[a]])[2]<- 
paste("rel",colnames(club1.real)[a],sep="_")
}

for (x in 2:13) {
  relative.club1.real[[x+1]]<-left_join(relative.club1.real[[x]],relative.club1.real[[x+1]], by="Province")
}
```


```{r}
table.relative.club1.real <- map(relative.club1.real, as.data.table)
df.relative.club1.real <- rbindlist(table.relative.club1.real, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.club1.real <- join(df.relative.club1.real,club1.real,by=c('Province')) %>% 
  select(Province,contains("rel"),Club)
  
df.relative.club1.real
```

```{r}
colnames(df.relative.club1.real) <- gsub("rel_","",colnames(df.relative.club1.real))
df.relative.club1.real
```

```{r}
df.relative.club1.real.long <- df.relative.club1.real %>%  pivot_longer(-c(Province,Club), names_to = "Time", values_to="Rel_Wage")
df.relative.club1.real.long
```

```{r}
df.relative.club1.real.path <- aggregate(Rel_Wage ~ Province + Time, df.relative.club1.real.long, mean) %>% 
  arrange(Province)
```

```{r}
path_club1.real <- df.relative.club1.real.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Province, col=Province)) + geom_line() +
labs(x = "Time",
y = "Relative transition paths of provinces in Club 1", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
path_club1.real
ggsave("path_club1.real.png", width = 9, height = 5)
```

#Club 2
```{r}
club2.real <- real.wage.club %>% 
  filter(Club == "Club 2")
```

```{r}
relative.club2.real <- list()
for(a in 2:14) {
  relative.club2.real[[a]] <- data.frame(club2.real$Province, club2.real[,a]/mean(club2.real[,a]))
colnames(relative.club2.real[[a]])[1]<- "Province"
colnames(relative.club2.real[[a]])[2]<- 
paste("rel",colnames(club2.real)[a],sep="_")
}

for (x in 2:13) {
  relative.club2.real[[x+1]]<-left_join(relative.club2.real[[x]],relative.club2.real[[x+1]], by="Province")
}
```


```{r}
table.relative.club2.real <- map(relative.club2.real, as.data.table)
df.relative.club2.real <- rbindlist(table.relative.club2.real, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.club2.real <- join(df.relative.club2.real,club2.real,by=c('Province')) %>% 
  select(Province,contains("rel"),Club)
  
df.relative.club2.real
```

```{r}
colnames(df.relative.club2.real) <- gsub("rel_","",colnames(df.relative.club2.real))
df.relative.club2.real
```

```{r}
df.relative.club2.real.long <- df.relative.club2.real %>%  pivot_longer(-c(Province,Club), names_to = "Time", values_to="Rel_Wage")
df.relative.club2.real.long
```

```{r}
df.relative.club2.real.path <- aggregate(Rel_Wage ~ Province + Time, df.relative.club2.real.long, mean) %>% 
  arrange(Province)
```

```{r}
path_club2.real <- df.relative.club2.real.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Province, col=Province)) + geom_line() +
labs(x = "Time",
y = "Relative transition paths of provinces in Club 2", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
path_club2.real
ggsave("path_club2.real.png", width = 9, height = 5)
```

#Club 3
```{r}
club3.real <- real.wage.club %>% 
  filter(Club == "Club 3")
```

```{r}
relative.club3.real <- list()
for(a in 2:14) {
  relative.club3.real[[a]] <- data.frame(club3.real$Province, club3.real[,a]/mean(club3.real[,a]))
colnames(relative.club3.real[[a]])[1]<- "Province"
colnames(relative.club3.real[[a]])[2]<- 
paste("rel",colnames(club3.real)[a],sep="_")
}

for (x in 2:13) {
  relative.club3.real[[x+1]]<-left_join(relative.club3.real[[x]],relative.club3.real[[x+1]], by="Province")
}
```


```{r}
table.relative.club3.real <- map(relative.club3.real, as.data.table)
df.relative.club3.real <- rbindlist(table.relative.club3.real, fill = T, idcol = T) %>% drop_na()
library(plyr)
df.relative.club3.real <- join(df.relative.club3.real,club3.real,by=c('Province')) %>% 
  select(Province,contains("rel"),Club)
  
df.relative.club3.real
```

```{r}
colnames(df.relative.club3.real) <- gsub("rel_","",colnames(df.relative.club3.real))
df.relative.club3.real
```

```{r}
df.relative.club3.real.long <- df.relative.club3.real %>%  pivot_longer(-c(Province,Club), names_to = "Time", values_to="Rel_Wage")
df.relative.club3.real.long
```

```{r}
df.relative.club3.real.path <- aggregate(Rel_Wage ~ Province + Time, df.relative.club3.real.long, mean) %>% 
  arrange(Province)
```

```{r}
path_club3.real <- df.relative.club3.real.path %>% 
  ggplot(aes(x=Time,y=Rel_Wage, group=Province, col=Province)) + geom_line() +
labs(x = "Time",
y = "Relative transition paths of provinces in Club 3", color="") +
theme(text=element_text( family="Palatino")) + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) 
  guides(col = guide_legend(nrow = 8))
path_club3.real
ggsave("path_club3.real.png", width = 9, height = 5)
```





# Logit model
# Load the data
```{r}
library(readxl)
logit <- read_excel("WS Labour_new.xlsx", 
    sheet = "logit")
logit
```

```{r}
logit$clubs <- factor(logit$clubs)
```

```{r}
logit <- logit %>% 
  mutate(lngrdp = log(grdp))
```

```{r}
glimpse(logit)
```

```{r}
mylogit <- glm(clubs ~ real_wage_2008 + lbr_prod_g +  lngrdp  + poverty, data = logit, family = "binomial")
summary(mylogit)
```